---
title: "ESS"
author: "Rebecca & Simon "
#output: html_notebook
output: github_document
---

## ToDo

* Build github Repo (Name?) -> social data science journey.
* Derive research question
* data wrangling
* fitting models
* write semester paper:
    + Die Studierenden führen für 12 LP abschließend eine eigene Untersuchung durch zu **zwei oder mehr selbst gewählten Beteiligungs- bzw. Entscheidungsprozessen**, um eine konkrete Forschungsfrage zu beantworten (ca. 30 Seiten). Wieviele Seiten zu Zweit?
* http://students.brown.edu/seeing-theory/basic-probability/index.html#section2

## Research Question(s)

* internal and external political efficacy  -> likelihood to vote right/populist or green party? Cultural clash. 
* Are citizens who live in systems that provide extensive direct democratic decision-making processes characterized by higher levels of trust in political institutions and higher external political efficacy?
* ESS in combination with direct democracy dataset?
* Macro predictors: polticial system or number of referenda
* Multilevel-Baysian Regression. 



## Packages

```{r}
pacman::p_load(dplyr, ggplot2, readr, haven, broom, purrr, tidyr, magrittr, labelled, sjPlot, viridis, forcats, ggthemes)
```

## Data

```{r}
ess <- read_sav("data/ESS8e01.sav")
ess_add <- read_sav("data/ESS1-7e01.sav") %>%
  filter(! cntry == "RU" & !cntry == 6)
# unique(ess$cntry) %in% unique(ess_add$cntry)
```

Select Variables from ESS round 8 and the cummulative round 1-7 for missing country data ([ESS Long Format](http://www.europeansocialsurvey.org/downloadwizard/)). The countries are augumented from: 

* `Round 6`: Ukraine, Slovakia, Hugary, Cyprus, Bulgaria, Albania
* `Round 5`: Turkey, Kosovo, Italy, Greece, Croatia
* `Round 7`: Spain, Protugal, Lithuania, Denmark

```{r}
tx_clean_spss <- function(x){
  x %>%     
    mutate_if(is.numeric, as.numeric) %>%
    mutate_if(is.character, as.character) %>%
    mutate_if(is.factor, as.factor)
}

get_party <- function(x){
  x %>% 
    select(contains("prtvt")) %>%
    mutate_all(sjmisc::to_label) %>% 
    mutate_all(as.character) %>%
    mutate(vote = apply(., 1, function(x) paste(ifelse(is.na(x), "", x), collapse = "")))
}

select_both <- function(x){
  x %>%
    select(idno, cntry, essround, hinctnta, polintr, lrscale,
         rlgatnd, pbldmn, ctzcntr, rlgatnd, edulvlb, eduyrs, yrbrn, gndr,
         dplyr::starts_with("trst"),
         dplyr::starts_with("stf"),
         #dplyr::starts_with("im"),
         imbgeco, imsmetn, imdfetn, impcntr,
         #dplyr::starts_with("prtvt")
    ) %>%
    mutate(cname = sjmisc::to_label(cntry)) %>% 
    mutate(cname = countrycode::countrycode(cname, "country.name", "country.name")) %>%
    mutate(iso2 = countrycode::countrycode(cname, "country.name", "iso2c")) %>%
    mutate(iso3 = countrycode::countrycode(cname, "country.name", "iso3c")) %>%
    rename(
      id = idno, 
      round = essround, 
      income = hinctnta,
      pol_inter = polintr,
      gndr = gndr,
      year = yrbrn,
      edu_year = eduyrs,
      edu = edulvlb,
      rel = rlgatnd,
      demo = pbldmn,
      citz = ctzcntr,
      trust_parl = trstprl, 
      trust_pol = trstplt,
      trust_party = trstprt,
      trust_leg = trstlgl,
      trust_police = trstplc,
      trust_eu = trstep,
      s_life = stflife,
      s_econ = stfeco,
      s_gov = stfgov,
      s_dem = stfdem,
      imm_econ = imbgeco,
      imm_same = imsmetn,
      imm_diff = imdfetn,
      imm_poor = impcntr
    )
}
```



```{r}
# spss data frame carry additional attribute information that is redudant and errorious for further analysis. 
ess_prep <- ess %>% 
  select_both()

vote <- ess %>%
  get_party()
ess_prep$vote <- vote$vote

ess_add_prep <- ess_add %>% 
  select_both()

vote <- ess_add %>%
  get_party()
ess_add_prep$vote <- vote$vote


sh <- c("idno",  "nwspol", "etusoft", "polintr", "sppsgva", "ctrolga", "sppipla", "ptppola", "wrkprty", "sgnptit",  "pbldmn", "lrscale", "stflife",  "stfeco",  "stfgov",  "stfdem",  "stfedu", "stfhlth", "euftf", "imbgeco", "imsmetn", "imdfetn", "impcntr", "rlgatnd", "ctzcntr", "sblazy", "gndr", "yrbrn", "eduyrs", "dulvlb", "inctnta", "region")

for(jj in sh){
  print(jj)
  print(any(colnames(ess_add) %in% jj))
}

### Not included so far
# news_pol = nwspol
# inet = netusoft
# poli_intr = polintr
# have_say = psppsgva
# pol_grp = actrolga
# have_infl = psppipla
# in_eff = cptppola
# party_act = wrkprty
# sign_pet = sgnptit
# state_edu = stfedu
# sate_health = stfhlth
# eu_state = euftf
# social_lazy = sblazy
# region = regio
```


Diverging variable naming due to country time space needs special management and renaming. 

```{r}
# ess_prep %>% map_lgl(~any(!is.na(.)))
# select which columns do not have NA only
vote_sel <- ess_add_prep %>%
  map_lgl(~any(!is.na(.)))

ess_add_prep <- ess_add_prep %>%
  select(which(vote_sel))

# test if NA remain in vote variables
ess_add_prep %>% map_lgl(~any(!is.na(.)))

# bidn round 8 and 5-7
ess_final <- bind_rows(ess_prep, ess_add_prep) %>%
  tx_clean_spss()

# rename voting columns
cols <- ess_final %>%
  select(contains("prtvt")) %>%
  colnames() %>%
  stringr::str_extract("..$") %>%
  paste0("vote_",.)

index <- which(stringr::str_detect(colnames(ess_final), "prtvt"))
colnames(ess_final)[index] <- cols

# ess_final %>%
#   group_by(cntry) %>%
#   slice(1)
# 
# ess_final %>% 
#   count(cntry, round) %>% 
#   count(cntry)
colnames(ess_final)
table(ess_final$vote)
glimpse(ess_final)


ess_final %>% 
  group_by(cntry) %>%
  count(is.na(vote))
```


## Country by Round

```{r}
ess_final %>% 
  count(cntry, round) %>%
  mutate(cntry = as.character(cntry)) %>%
  mutate(round = as.factor(round)) %>%
  mutate(cntry = forcats::fct_reorder(cntry, n)) %>%
  ggplot(aes(cntry, n, fill = round, colour = round, label = n)) +
  geom_bar(stat = "identity", alpha = .3, colour = NA) +
  geom_point(size = 3) + 
  coord_flip() +
  #ylim(800, 3000) +
  geom_text(nudge_y = 150) +
  viridis::scale_fill_viridis(discrete = T, direction = -1) +
  viridis::scale_colour_viridis(discrete = T, direction = -1) +
  theme_hc()

#plotly::ggplotly()
```

```{r}
glimpse(ess_final)
```

## Map Countries Selected

```{r}
library(ggplot2)
world <- map_data("world")
world$iso3 <- countrycode::countrycode(world$region, "country.name", "iso3c")
world$value <- ifelse(world$iso3 %in% unique(ess_final$iso3), "yes", "no")
# table(world$value)
# world %>% 
#   ggplot(aes(long, lat, group = group)) + 
#     geom_polygon(fill='grey')

world %>% 
  ggplot(aes(long, lat, group = group, fill = value)) + 
  geom_polygon() +
  xlim(-20,50) + 
  ylim(30,80) +
  scale_fill_manual("selected", values = c("gray90", "blue")) +
  theme_map()
```



## Match party data 


```{r}
ess_final %<>% 
  mutate(populism = case_when(

    #Belgium
    vote == "Parti Populaire" ~ "Illiberal Populism",
    vote == "PVDA+" ~ "Liberal Populism",
    vote == "Vlaams Belang" ~ "Illiberal Populism",
    
    #Denmark
    vote == "Dansk Folkeparti - Danish peoples party" ~ "Illiberal Populism",
    vote == "Enhedslisten - Unity List - The Red-Green Alliance" ~ "Liberal Populism",
    #Folkebevægelsen mod EU is missing   
    
    #Deutschland
    vote == "NPD" ~ "Illiberal Populism",
    vote == "AfD" ~ "Illiberal Populism",
    vote == "Die Linke" ~ "Liberal Populism",
    vote == "Piratenpartei" ~ "Liberal Populism",
    #Die Tierpartei not included
    
    #Greece
    vote == "Golden Dawn (Xrusi Augi)" ~ "Illiberal Populism",
    vote == "The Coalition of the Radical Left (SYRIZA)" ~ "Liberal Populism",
    vote == "The Communist Party of Greece (KKE)" ~ "Illiberal Populism",
    vote == "The Popular Orthodox Rally (LAOS)" ~ "Illiberal Populism",
    #vote == "Independent Citizens' Movement (SPITHA)" ~ "Illiberal Populism",
    #ANEL gab es 2010 noch gar nicht :(

    #Spain
    vote == "AMAIUR" ~ "Liberal Populism",
    vote == "Bloque Nacionalista Galego (BNG)" ~ "Liberal Populism",
    vote == "Izquierda Unida (IU) - (ICV en Cataluña)" ~ "Liberal Populism",
    #PODEMOS was created in 2014
  
    #Frankreich
    vote == "FN (Front National)" ~ "Illiberal Populism",
    vote == "MPF (Mouvement pour la France)" ~ "Illiberal Populism",
    vote == "FDG (Front de Gauche)" ~ "Liberal Populism",
    #fr_PCF	Parti Communiste Français	
    #fr_PG	Parti de Gauche	
    #fr_Ensemble	Ensemble  --> unter FDG
    #PG, MPF, PCF und Ensamble fehlt

    #Ireland
    vote == "Green Party" ~ "Liberal Populism",
    vote == "Sinn Féin" ~ "Liberal Populism",
    vote == "Anti-Austerity Alliance - People Before Profit" ~ "Liberal Populism",
    vote == "Socialist Party - United Left Alliance" ~ "Liberal Populism",

    #Italy
    vote == "Movimento 5 Stelle" ~ "Liberal Populism",
    vote == "Fratelli d'Italia" ~ "Illiberal Populism",
    vote == "Lega Nord" ~ "Illiberal Populism",
    vote == "Sinistra Ecologia e Libertà (SEL)" ~ "Liberal Populism",
    #it_RC	Partito della Rifondazione Comunista	missing

    #Netherland
    vote == "Socialist Party" ~ "Liberal Populism",
    vote == "Party for Freedom" ~ "Illiberal Populism", 
    vote == "Party for the Animals" ~ "Liberal Populism",
    vote == "50PLUS" ~ "Liberal Populism",

    #UK
    vote == "Green Party" ~ "Liberal Populism",
    vote == "UK Independence Party" ~ "Illiberal Populism",

    #Portugal
    vote == "Bloco de Esquerda (BE)" ~ "Liberal Populism",
    vote == "Coligação Democrática Unitária (CDU)" ~ "Liberal Populism",
    #por_MPT	Partido da Terra missing

    #Österreich
    vote == "BZÖ" ~ "Illiberal Populism",
    vote == "FPÖ" ~ "Illiberal Populism",    
    vote == "Team Frank Stronach" ~ "Illiberal Populism",  

    #Finland  
    vote == "True Finns" ~ "Illiberal Populism",
    vote == "Left Alliance" ~ "Liberal Populism",

    #Sweden
    vote == "FI (Feministiskt initiativ)" ~ "Liberal Populism",
    vote == "Piratpartiet" ~ "Liberal Populism", 
#    vote == "Sverigedomkraterna" ~ "Illiberal Populism", 
    stringr::str_detect(vote, pattern = "Sverig") ~ "Illiberal Populism",  
    vote == "Vänsterpartiet" ~ "Liberal Populism",

    #Bulgaria
    vote == "PP Ataka" ~ "Illiberal Populism",
    vote == "PP VMRO - Balgarsko natsionalno dvizhenie" ~ "Illiberal Populism", 
    vote == "PP Natsionalen front za dpasenie na Bulgaria" ~ "Illiberal Populism", 
    #bul_BBT	Bulgaria bez Tsenzura wasn't there in 2014

    #Czech Republic
    vote == "KSČM" ~ "Illiberal Populism",
    vote == "ANO 2011" ~ "Liberal Populism",
    vote == "Úsvit přímé demokracie Tomia Okamury" ~ "Illiberal Populism", 
    #cz_SVOBODNI.... miss you
      
    #Estonia
    vote == "Erakond Eestimaa Rohelised" ~ "Liberal Populism",
    vote == "Eesti Vabaerakond" ~ "Liberal Populism",

    #Hungary
    #vote == "Fidesz - KDNP (Fidesz – Magyar Polgári Szövetség Keresztényd" ~ "Illiberal Populism",
    stringr::str_detect(vote, "Fidesz") ~ "Illiberal Populism",
    vote == "Jobbik (Jobbik Magyarországért Mozgalom)" ~ "Illiberal Populism",
    vote == "LMP (Lehet Más A Politika)" ~ "Liberal Populism",


    #Lithuania
    vote == "Electoral Action of Poles in Lithuania (LLRA)" ~ "Illiberal Populism",
    vote == "Party Order and Justice (TT)" ~ "Illiberal Populism",
    vote == "Political Party 'The Way of Courage' (DK)" ~ "Illiberal Populism",


    #Poland
    vote == "Law and Justice" ~ "Illiberal Populism",
    vote == "Together Party" ~ "Illiberal Populism",
    vote == "Kukiz'15" ~ "Illiberal Populism",
    #pol_KNP	Kongres Nowej Prawicy könnte auch Kukiz'15 sein
    

    #Slovakia; 
    vote == "Freedom and Solidarity (SaS)" ~ "Liberal Populism",
    vote == "Ordinary People and Independent Personalities (OLaNO)" ~ "Illiberal Populism",
    #slo_SNS	Slovenská národná strana	 Missing
    
    #Croatia; 
    vote == "Hrvatska stranka prava (HSP)" ~ "Illiberal Populism",


    #Cyprus; 
    vote == "Progressive Party of Working People (AKEL)" ~ "Liberal Populism",
    vote == "The Cyprus Green Party" ~ "Liberal Populism",
    
    is.na(vote) ~ NA_character_,
    TRUE ~ "Establishment"
   ))

ess_final %>% 
  group_by(cntry, populism) %>%
  tally %>%
  count(cntry)
```





## binoculaR

`binoculaR` is an own product developed to mimic SPSS in order to quickly inspect variables and their corresponding levels/ attributes. 

```{r, eval = F}
#devtools::install_github("systats/binoculaR", force = T)
selected <- binoculaR::binoculaR(ess)
save(selected, file = "selected.Rdata")
```


List all variables (old and new names) as well as their range of levels. 


```{r, eval = F}
ess_clean <- ess_sub  %>% 
  # mutate(eu_member =
  #          recode_factor(cntry,
  #               DE = 1958, BE = 1958, FR = 1958, NL = 1958, IE = 1973,
  #               GB = 1973, FI = 1995, AT = 1995, SE = 1995, EE = 2004,
  #               PL = 2004, SI = 2004, CZ = 2004, CH = 0, IL = 0,
  #               IS = 0, NO = 0, RU = 0
  #             )
  #       ) %>%
  # mutate(post_com = ifelse(region %in% c("Estonia", "Poland", "Slovenia", "Czech Republic", "Russian Federation"), "Post C", "West"))
```

## Get Info

* better left-right scale more athoritarian
* horse shoe theory
* Set up Rmarkdown paper template
* Set up Project page
* Paraboost [D3partitionR](https://github.com/AntoineGuillot2/D3partitionR)


## Data 

* party recoding
* Inspect each variable with a histogram or barplot, check range and validate data. 


### Block 5

* clean Variables
* document range

```{r, eval = F}
p5 <- ess_clean %>%
  select(gndr, edu, income, rel, year) %>%
  gather("var", "value") %>%
  ggplot(aes(value, fill = var)) +
  geom_bar() +
  facet_wrap(~var, scales = "free") +
  viridis::scale_fill_viridis(discrete = T)
p5
```





## EDA

* Check selective or pairwise compare correlations (stepwiseAIC or multilevel predction tree).
* spatial mapping
* Factor scores (Factor Analysis)
    + Trust (target)
    + Imm (X)

```{r eda1, eval = F,  echo = F, fig.height=6, fig.width=10, results="show"}
# citizens trust in european institutions 
d <- ess_sub

# graph settings
set_theme(#geom.outline.color = "antiquewhite4", 
          #geom.outline.size = 1, 
          #geom.label.size = 2,
          #geom.label.color = "grey50",
          title.color = "black", 
          title.size = 1.5, 
          axis.angle.x = 45, 
          axis.textcolor = "black", 
          base = theme_538())

p1 <- sjp.frq(d$trust_parl, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p2 <- sjp.frq(d$trust_leg, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p3 <- sjp.frq(d$trust_police, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p4 <- sjp.frq(d$trust_parl, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p5 <- sjp.frq(d$trust_party, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p6 <- sjp.frq(d$trust_pol, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

library(gridExtra)
grid.arrange(p1$plot, p2$plot, p3$plot, p4$plot, p5$plot, p6$plot, nrow = 3, top = grid::textGrob("Trust in Political Institutions", vjust = 0.5, gp = grid::gpar(fontface = "bold", cex = 2)))
```




## Statistical Modeling

### Multilevel 

* Y = voting right wing/populist
* X_1 = ex_eff (trust factor scores) + in_eff
* X_2 = immigration(factor scores)
* X_3 = s_gov + s_dem + interaction + demo + sign_pet
* X_4 = poli_intr + lrscale + party_act + news_pol
* X_5 = gndr + edu + income + rel + year
* X_6 = social_lazy (social benefit attitudes)
* What about Z predictors?


### Prection Power

* multilevel tree
* `brms`
* viele random slopes
