---
title: "ESS Data Wrangling"
subtitle: "Testing differnt causal paths..."
author: "Rebecca & Simon "
output: html_notebook
---

## Packages

```{r}
pacman::p_load(dplyr, ggplot2, readr, haven, broom, purrr, tidyr, magrittr, labelled, sjPlot, viridis, forcats, ggthemes)
```

## Data

Select Variables from ESS round 8 and the cummulative round 1-7 for missing country data ([ESS Long Format](http://www.europeansocialsurvey.org/downloadwizard/)). The countries are augumented from: 

* `Round 6`: Ukraine, Slovakia, Hugary, Cyprus, Bulgaria, Albania
* `Round 5`: Turkey, Kosovo, Italy, Greece, Croatia
* `Round 7`: Spain, Protugal, Lithuania, Denmark

```{r, warning=F, message=F, error=F, results="hide"}
get_party <- function(x){
  dat1 <- x %>% 
    select(contains("prtv"), prtvede1, prtvede2) %>%
    mutate_all(sjmisc::to_label) %>% 
    mutate_all(as.character) %>%
    mutate(vote = apply(., 1, function(x) paste(ifelse(is.na(x), "", x), collapse = ""))) %>%
    mutate(vote = vote %>%
           stringr::str_replace_all("SPD", " SPD ") %>%
           stringr::str_replace_all("CDU/CSU", " CDU/CSU ") %>%
           stringr::str_replace_all("FDP", " FDP ") %>%
           stringr::str_replace_all("NPD", " NPD ") %>%
           stringr::str_replace_all("Die Linke", " DieLinke ") %>%
           stringr::str_replace_all("AfD", " AfD ") %>%
           stringr::str_replace_all("Andere Partei", " AnderePartei ") %>%
           stringr::str_replace_all("Piratenpartei", " Piratenpartei ") %>%
           stringr::str_replace_all("Bündnis.*?Grünen", " Bündnis90/DieGrünen ") %>%
           stringr::str_replace_all("^Grünen", " Bündnis90/DieGrünen ") %>%
           stringr::str_replace_all("\\s+", " ") %>%
           stringr::str_trim() %>%
           stringr::str_replace_all("^.*? ", "")
         )
  
  vote_id <- x %>%
    select(contains("prtv"), prtvede1, prtvede2) %>%
    #mutate_all(sjmisc::to_label) %>% 
    mutate_all(as.character) %>%
    split(1:nrow(x)) %>%
    map_chr(~ paste(ifelse(is.na(.x), "", .x), collapse = ""))

  x %>%
    mutate(vote = dat1$vote, vote_id = vote_id)
}

ess <- read_sav("data/original/ESS8e01.sav") %>%
  mutate(cname = sjmisc::to_label(cntry)) %>%
  get_party()

ess_add <- read_sav("data/original/ESS1-7e01.sav") %>%
  filter(! cntry == "RU" & !cntry == 6) %>%
  mutate(cname = sjmisc::to_label(cntry)) %>%
  get_party()
# unique(ess$cntry) %in% unique(ess_add$cntry)

ess_all <- bind_rows(ess, ess_add)
dim(ess_all)

step1 <- ess_all %>%
    select(idno, cntry, cname, essround, hinctnta, polintr, lrscale,
         rlgatnd, pbldmn, ctzcntr, rlgatnd, edulvlb, eduyrs, yrbrn, gndr,
         dplyr::starts_with("trst"),
         dplyr::starts_with("stf"),
         #dplyr::starts_with("im"),
         imbgeco, imsmetn, imdfetn, impcntr,
         #dplyr::starts_with("prtvt")
         vote, vote_id
    )

table(step1$vote_id)
```


```{r}
step2 <- step1 %>%
    mutate(country = countrycode::countrycode(cname, "country.name", "country.name")) %>%
    mutate(iso2 = countrycode::countrycode(country, "country.name", "iso2c")) %>%
    mutate(iso3 = countrycode::countrycode(cname, "country.name", "iso3c")) %>%
    select(-cntry, -cname)
colnames(step2)
```


```{r}
step3 <- step2 %>%
    rename(
      id = idno, 
      round = essround, 
      income = hinctnta,
      pol_inter = polintr,
      gndr = gndr,
      year = yrbrn,
      edu_year = eduyrs,
      edu = edulvlb,
      rel = rlgatnd,
      demo = pbldmn,
      citz = ctzcntr,
      trust_parl = trstprl, 
      trust_pol = trstplt,
      trust_party = trstprt,
      trust_leg = trstlgl,
      trust_police = trstplc,
      trust_eu = trstep,
      s_life = stflife,
      s_econ = stfeco,
      s_gov = stfgov,
      s_dem = stfdem,
      imm_econ = imbgeco,
      imm_same = imsmetn,
      imm_diff = imdfetn,
      imm_poor = impcntr
    )

### Not included so far
# news_pol = nwspol
# inet = netusoft
# poli_intr = polintr
# have_say = psppsgva
# pol_grp = actrolga
# have_infl = psppipla
# in_eff = cptppola
# party_act = wrkprty
# sign_pet = sgnptit
# state_edu = stfedu
# sate_health = stfhlth
# eu_state = euftf
# social_lazy = sblazy
# region = regio

glimpse(step3)
```

```{r}
### Germany special 
step3$vote_id[step3$country == "Germany"] <- step3 %>%
  filter(country == "Germany") %>%
  mutate(vote_id = stringr::str_extract(vote_id, "\\d$")) %>%
  .$vote_id
```

Filter problematic countries

```{r}
step4 <- step3 %>%
  filter(!(country %in% c("Romania", "Malta", "Luxembourg", "Latvia", "Lithuania")))%>%
  mutate(vote_id = ifelse(vote_id == "", "", paste(iso2, vote_id, sep = "_")))
```


```{r}
ess_parties <- step4 %>%
  filter(vote != "") %>%
  group_by(country, iso2, vote, vote_id) %>%
  tally %>%
  #mutate(n = NULL) %>%
  filter(!vote == "Other") %>%
  arrange(country, vote_id) %>%
  mutate(vote_id = paste(iso2, vote_id, sep = "_"))


library(openxlsx)
#openxlsx::write.xlsx(ess_parties, file = "data/ess_parties_new.xlsx")
# save(ess_parties, file = "shiny_match/ess_parties.Rdata")
# write.csv2(ess_parties, "data/ess_parties.csv")
matching_data <- openxlsx::read.xlsx("data/original/ess_parties_new.xlsx") %>%
  select(party_id, vote_id) %>%
  na.omit
```

## Match party data 

mapping ches data to ess

```{r}
glimpse(step4)

ess_final <- step4 %>%
  left_join(matching_data, by = "vote_id")

# save(ess_final, file = "data/Rdata/ess_final.Rdata")

ess_final %>%
  filter(vote != "") %>%
  count(is.na(party_id))
```


# CHES

## PDF to dataframe

```{r}
# pacman::p_load(pdftables)
# 
# pdftable_api <- "opj9i5owyg40"
# pdftables::convert_pdf("data/2014_CHES_codebook.pdf",
#                         "data/party_tables.csv",
#                         api_key = pdftable_api)
# 
# pdftables::convert_pdf("codebooks/country_names.pdf",
#                         "codebooks/country_names.csv",
#                         api_key = pdftable_api)
# 
```

## Import and ESS matching

```{r}

clist <- list(
  BE = "Belgium",
  DK  = "Denmark",
  GE  = "Germany",
  GR  = "Greece",
  ESP = "Spain",
  FR  = "France",
  IRL = "Ireland",
  IT  = "Italy",
  NL = "Netherlands",
  UK = "United Kingdom",
  POR=  "Portugal",
  AUS=  "Austria",
  FIN=  "Finland",
  SV = "Sweden",
  BUL=  "Bulgaria",
  CZ = "Czech Republic",
  EST=  "Estonia",
  HUN=  "Hungary",
  LAT=  "Latvia",
  LITH = "Lithuania",
  POL=  "Poland",
  ROM=  "Romania",
  SLO=  "Slovakia",
  SLE=  "Slovenia",
  CRO=  "Croatia",
  TUR=  "Turkey",
  NOR=  "Norway",
  SWI=  "Switzerland",
  MAL=  "Malta",
  LUX=  "Luxembourg",
  CYP=  "Cyprus"
)

library(haven)
ches <- read_dta("data/original/2014_CHES_dataset_means-2.dta") %>%
  mutate(country = tidyTX::tx_map_dict(country, dict = clist, key1 = 0, key2 = 1)) %>%
  mutate(party_id = as.integer(party_id))

ches_final <- ches %>%
  left_join(matching_data, by = "party_id")

# save(ches_final, file = "data/Rdata/ches_final.Rdata")

# dt <- read_csv("data/party_tables.csv") %>%
#   mutate(country = tidyTX::tx_map_dict(country, dict = clist, key1 = 0, key2 = 1)) %>%
#   mutate(party_id = as.integer(party_id))

# ches_data <- ches %>%
#   select(-country, -party_name) %>%
#   mutate(party_id = as.integer(party_id)) %>%
#   left_join(dt, by = c("party_id")) %>%
#   mutate(country = countrycode::countrycode(country, "country.name", "country.name"))

# ches_parties <- ches_data %>% 
#   select(country, party_id, party_short, party_name_en) %>%
#   mutate(vote_id = NA) %>%
#   arrange(country)
# 
# table(ches_data$country)
# #ches_party <- nn %>%
# # save(ches_parties, file = "shiny_match/ches_parties.Rdata")
# #write.csv2(ches_parties, "data/ches_parties.csv")
# library(openxlsx)
# #openxlsx::write.xlsx(ches_parties, file = "data/ches_parties.xlsx")
# #readr::write_csv(ches_parties, path = "data/ches_parties.csv")
```