---
title: "ESS Data Wrangling"
subtitle: "Testing differnt causal paths..."
author: "Rebecca & Simon "
output: html_notebook
---

## Packages

```{r}
pacman::p_load(dplyr, ggplot2, readr, haven, broom, purrr, tidyr, magrittr, labelled, sjPlot, viridis, forcats, ggthemes)
```

## Data

Select Variables from ESS round 8 and the cummulative round 1-7 for missing country data ([ESS Long Format](http://www.europeansocialsurvey.org/downloadwizard/)). The countries are augumented from: 

* `Round 6`: Ukraine, Slovakia, Hugary, Cyprus, Bulgaria, Albania
* `Round 5`: Turkey, Kosovo, Italy, Greece, Croatia
* `Round 7`: Spain, Protugal, Lithuania, Denmark


```{r, warning=F, message=F, error=F, results="hide"}
get_party <- function(x){
  dat1 <- x %>% 
    select(contains("prtv"), prtvede1, prtvede2) %>%
    mutate_all(sjmisc::to_label) %>% 
    mutate_all(as.character) %>%
    mutate(vote = apply(., 1, function(x) paste(ifelse(is.na(x), "", x), collapse = ""))) %>%
    mutate(vote = vote %>%
           stringr::str_replace_all("SPD", " SPD ") %>%
           stringr::str_replace_all("CDU/CSU", " CDU/CSU ") %>%
           stringr::str_replace_all("FDP", " FDP ") %>%
           stringr::str_replace_all("NPD", " NPD ") %>%
           stringr::str_replace_all("Die Linke", " DieLinke ") %>%
           stringr::str_replace_all("AfD", " AfD ") %>%
           stringr::str_replace_all("Andere Partei", " AnderePartei ") %>%
           stringr::str_replace_all("Piratenpartei", " Piratenpartei ") %>%
           stringr::str_replace_all("Bündnis.*?Grünen", " Bündnis90/DieGrünen ") %>%
           stringr::str_replace_all("^Grünen", " Bündnis90/DieGrünen ") %>%
           stringr::str_replace_all("\\s+", " ") %>%
           stringr::str_trim() %>%
           stringr::str_replace_all("^.*? ", "")
         )
  
  vote_id <- x %>%
    select(contains("prtv"), prtvede1, prtvede2) %>%
    #mutate_all(sjmisc::to_label) %>% 
    mutate_all(as.character) %>%
    split(1:nrow(x)) %>%
    map_chr(~ paste(ifelse(is.na(.x), "", .x), collapse = ""))

  x %>%
    mutate(vote = dat1$vote, vote_id = vote_id)
}

ess <- read_sav("data/original/ESS8e01.sav") %>%
  mutate(cname = sjmisc::to_label(cntry)) %>%
  get_party()

ess_add <- read_sav("data/original/ESS1-7e01.sav") %>%
  filter(! cntry == "RU" & !cntry == 6) %>%
  mutate(cname = sjmisc::to_label(cntry)) %>%
  get_party()
# unique(ess$cntry) %in% unique(ess_add$cntry)

ess_all <- bind_rows(ess, ess_add)
dim(ess_all)




step1 <- ess_all %>%
    select(idno, cntry, cname, essround, region, gndr, yrbrn, hinctnta, polintr,
         lrscale, eduyrs, edulvlb, rlgatnd, psppsgva, psppipla, actrolga, cptppola, 
         dplyr::starts_with("trst"),
         stflife, stfeco, stfgov, stfdem, stfedu, stfhlth,
         #dplyr::starts_with("im"),
         imbgeco, imsmetn, imdfetn, impcntr,
         #dplyr::starts_with("prtvt")
         vote, vote_id
    ) %>%
    mutate(country = countrycode::countrycode(cname, "country.name", "country.name")) %>%
    mutate(iso2 = countrycode::countrycode(country, "country.name", "iso2c")) %>%
    mutate(iso3 = countrycode::countrycode(cname, "country.name", "iso3c")) %>%
    select(-cntry, -cname)

colnames(step1)
```


List all variables (old and new names) as well as their range of levels. 


**ID and Individual Characteristica**

* `id` (`idno`):
* `country` (`cntry`):
* `region` (`region`): NUTS
* `gender` (`gndr`): 1 Male - 2 Female
* `age` (`yrbrn`): 
* `edu_year` (`eduyrs`): full-time education in years
* `income` (`hinctnta`): Household's total net income 1 (1st decentile) - 10 (10th decentile)
* `poli_intr` (`polintr`):
* `lrscale` (`lrscale`): In politics people sometimes talk of "left" and "right". Using this card, where would you place yourself on this scale, where 0 means the left and 10 means the right? 1 (Extremely dissatisfied) - 10 (Extremely satisfied)
* `rel` (`rlgatnd`): Apart from special occasions such as weddings and funerals, about how often do you attend religious services nowadays? 1 (Every day) - 7 (Never)

**Trust**

0 means you do not trust an institution at all, and 10 means you have complete trust.


**EXT EFFICACY**

* `have_say` (`psppsgva`): How much would you say the political system in [country] allows people like you to have a say in what the government does? 1 (Not at all) - 5 (A great deal)
* `have_infl` (`psppipla`): And how much would you say that the political system in [country] allows people like you to have an influence on politics? 1 (Not at all) - 5 (A great deal)


**INT EFFICACY**

* `pol_grp` (`actrolga`): How able do you think you are to take an active role in a group involved with political issues? 1 (Not at all able) - 5 (Completely able) 
* `pol_conf` (`cptppola`): And how confident are you in your own ability to participate in politics? 1 (Not at all confident) - 5 (Completely confident )


**Satisfaction**

* `s_life` (`stflife`): All things considered, how satisfied are you with your life as a whole nowadays? Please answer using this card, where 0 means extremely dissatisfied and 10 means extremely satisfied. 1 (Extremely dissatisfied) - 10 (Extremely satisfied)
* `s_econ` (`stfeco`): On the whole how satisfied are you with the present state of the economy in [country]? 1 (Extremely dissatisfied) - 10 (Extremely satisfied)
* `s_gov` (`stfgov`): Now thinking about the [country] government, how satisfied are you with the way it is doing its job? 1 (Extremely dissatisfied) - 10 (Extremely satisfied)
* `s_dem` (`stfdem`): How satisfied with the way democracy works in country? 1 (Extremely dissatisfied) - 10 (Extremely satisfied)
* `state_edu` (`stfedu`): Now, using this card, please say what you think overall about the state of education in [country] nowadays?
* `sate_health` (`stfhlth`): Still using this card, please say what you think overall about the state of health services in [country] nowadays? 

**Cultural Deprivation**

* `imm_econ` (`imbgeco`): Would you say it is generally bad or good for [country]'s economy that people come to live here from other countries? 1 (Bad for the economy) - 10 (Good for the economy)
* `imm_same` (`imsmetn`): Now, using this card, to what extent do you think [country] should allow people of the same race or ethnic group as most [country] people to come and live here ? 1 (Allow many to come and live here) - 4 (Allow none)
* `imm_diff` (`imdfetn`): How about people of a different race or ethnic group from most [country] people? 1 (Allow many to come and live here) - 4 (Allow none) 
* `imm_poor` (`impcntr`): How about people from the poorer countries outside Europe? 1 (Allow many to come and live here) - 4 (Allow none)



```{r}
step2 <- step1 %>%
    rename(
      id = idno, 
      round = essround, 
      income = hinctnta,
      pol_inter = polintr,
      gndr = gndr,
      age = yrbrn,
      edu_year = eduyrs,
      edu = edulvlb,
      rel = rlgatnd,
      trust_parl = trstprl, 
      trust_pol = trstplt,
      trust_party = trstprt,
      trust_leg = trstlgl,
      trust_police = trstplc,
      trust_eu = trstep,
      s_life = stflife,
      s_econ = stfeco,
      s_gov = stfgov,
      s_dem = stfdem,
      state_edu = stfedu,
      sate_health = stfhlth,
      imm_econ = imbgeco,
      imm_same = imsmetn,
      imm_diff = imdfetn,
      imm_poor = impcntr,
      have_say = psppsgva,
      have_infl = psppipla,
      pol_grp = actrolga,
      pol_conf = cptppola
    )

glimpse(step2)
```

```{r}
### Germany special 
step2$vote_id[step2$country == "Germany"] <- step2 %>%
  filter(country == "Germany") %>%
  mutate(vote_id = stringr::str_extract(vote_id, "\\d$")) %>%
  .$vote_id
```

Filter problematic countries

```{r}
step3 <- step2 %>%
  filter(!(country %in% c("Romania", "Malta", "Luxembourg", "Latvia", "Lithuania")))%>%
  mutate(vote_id = ifelse(vote_id == "", "", paste(iso2, vote_id, sep = "_")))
```


```{r}
ess_parties <- step3 %>%
  filter(vote != "") %>%
  group_by(country, iso2, vote, vote_id) %>%
  tally %>%
  #mutate(n = NULL) %>%
  filter(!vote == "Other") %>%
  arrange(country, vote_id) %>%
  mutate(vote_id = paste(iso2, vote_id, sep = "_"))


library(openxlsx)
#openxlsx::write.xlsx(ess_parties, file = "data/ess_parties_new.xlsx")
# save(ess_parties, file = "shiny_match/ess_parties.Rdata")
# write.csv2(ess_parties, "data/ess_parties.csv")
matching_data <- openxlsx::read.xlsx("data/original/ess_parties_new.xlsx") %>%
  select(party_id, vote_id) %>%
  na.omit
```


## Match party data 

mapping ches data to ess

```{r}
ess_final <- step3 %>%
  left_join(matching_data, by = "vote_id")
glimpse(ess_final)
# save(ess_final, file = "data/Rdata/ess_final.Rdata")

ess_final %>%
  filter(vote != "") %>%
  count(is.na(party_id))
```


## CHES

```{r}
# pacman::p_load(pdftables)
# 
# pdftable_api <- "opj9i5owyg40"
# pdftables::convert_pdf("data/2014_CHES_codebook.pdf",
#                         "data/party_tables.csv",
#                         api_key = pdftable_api)
# 
# pdftables::convert_pdf("codebooks/country_names.pdf",
#                         "codebooks/country_names.csv",
#                         api_key = pdftable_api)
# 
```


Import and ESS matching

```{r}
clist <- list(
  BE = "Belgium",
  DK  = "Denmark",
  GE  = "Germany",
  GR  = "Greece",
  ESP = "Spain",
  FR  = "France",
  IRL = "Ireland",
  IT  = "Italy",
  NL = "Netherlands",
  UK = "United Kingdom",
  POR=  "Portugal",
  AUS=  "Austria",
  FIN=  "Finland",
  SV = "Sweden",
  BUL=  "Bulgaria",
  CZ = "Czech Republic",
  EST=  "Estonia",
  HUN=  "Hungary",
  LAT=  "Latvia",
  LITH = "Lithuania",
  POL=  "Poland",
  ROM=  "Romania",
  SLO=  "Slovakia",
  SLE=  "Slovenia",
  CRO=  "Croatia",
  TUR=  "Turkey",
  NOR=  "Norway",
  SWI=  "Switzerland",
  MAL=  "Malta",
  LUX=  "Luxembourg",
  CYP=  "Cyprus"
)

library(haven)
ches <- read_dta("data/original/2014_CHES_dataset_means-2.dta") %>%
  mutate(country = tidyTX::tx_map_dict(country, dict = clist, key1 = 0, key2 = 1)) %>%
  mutate(party_id = as.integer(party_id))

ches_final <- ches %>%
  left_join(matching_data, by = "party_id")

# save(ches_final, file = "data/Rdata/ches_final.Rdata")

# dt <- read_csv("data/party_tables.csv") %>%
#   mutate(country = tidyTX::tx_map_dict(country, dict = clist, key1 = 0, key2 = 1)) %>%
#   mutate(party_id = as.integer(party_id))

# ches_data <- ches %>%
#   select(-country, -party_name) %>%
#   mutate(party_id = as.integer(party_id)) %>%
#   left_join(dt, by = c("party_id")) %>%
#   mutate(country = countrycode::countrycode(country, "country.name", "country.name"))

# ches_parties <- ches_data %>% 
#   select(country, party_id, party_short, party_name_en) %>%
#   mutate(vote_id = NA) %>%
#   arrange(country)
# 
# table(ches_data$country)
# #ches_party <- nn %>%
# # save(ches_parties, file = "shiny_match/ches_parties.Rdata")
# #write.csv2(ches_parties, "data/ches_parties.csv")
# library(openxlsx)
# #openxlsx::write.xlsx(ches_parties, file = "data/ches_parties.xlsx")
# #readr::write_csv(ches_parties, path = "data/ches_parties.csv")
```


## V-Dem 

* `ori` = `v2xdd_i_or`: Obligatory referendum index
* Population size (E) (e_population)
Question: What is the size of the population?
* `pop_total` = `e_mipopula`: What is the total population (in millions)?
Clarifications: Missing data within a time-series is interpolated using a linear model,
increasing the number of observations from 3969 to 20979.

* `range_consul` = (`v2dlconslt`) Range of consultation: When important policy changes are being considered, how wide is the range
of consultation at elite levels?
    + 0: No consultation. The leader or a very small group (e.g. military council) makes authoritative decisions on their own.
    + 1: Very little and narrow. Consultation with only a narrow circle of loyal party/ruling elites.
    + 2: Consultation includes the former plus a larger group that is loyal to the government, such as the ruling party’s or parties’ local executives and/or women, youth and other branches.
    + 3: Consultation includes the former plus leaders of other parties.
    + 4: Consultation includes the former plus a select range of society/labor/business representatives.
    + 5: Consultation engages elites from essentially all parts of the political spectrum and all politically relevant sectors of society and business.

*  `engage_soc` = (`v2dlengage`) Engaged society: When important policy changes are being considered, how wide and how independent are public deliberations?
    + 0: Public deliberation is never, or almost never allowed.
    + 1: Some limited public deliberations are allowed but the public below the elite levels  almost always either unaware of major policy debates or unable to take part in them.
    + 2: Public deliberation is not repressed but nevertheless infrequent and non-elite tors are typically controlled and/or constrained by the elites.
    + 3: Public deliberation is actively encouraged and some autonomous non-elite groups rticipate, but it is confined to a small slice of specialized groups that tends to be e same across issue-areas.
    + 4: Public deliberation is actively encouraged and a relatively broad segment of n-elite groups often participate and vary with different issue-areas.
    + 5: Large numbers of non-elite groups as well as ordinary people tend to discuss major licies among themselves, in the media, in associations or neighborhoods, or in the reets. Grass-roots deliberation is common and unconstrained.

* ~~`perc_imm` = (`v2elnoncit`): Resident noncitizens who cannot vote (C).  What percentage (%) of the people residing in the country (or colony) does not enjoy the legal right to vote in national elections because they are not full citizens?~~



```{r}
# devtools::install_github("xmarquez/vdem")
# library(vdem)
vdem <- read.csv("data/original/vdem7.1/vdem.csv")

vdem_final <- vdem %>%
  select(country_name, year, v2xdd_i_or, v2dlconslt, v2dlengage) %>%
  rename(ori = v2xdd_i_or, range_consul = v2dlconslt, engage_soc = v2dlengage, country = country_name)
```


## Human Development Index

http://hdr.undp.org/en/content/human-development-index-hdi
http://hdr.undp.org/en/data

```{r}
dir("data/original")

hdi_final <- read.csv("data/original/hdi.csv", skip = 1, stringsAsFactors = F) %>%
  select(-contains(".")) %>%
  gather("year", "hdi", -Country) %>%
  mutate(year = year %>% stringr::str_extract("\\d{4}")) %>%
  rename(country = Country) %>% 
  mutate(year = as.character(year)) %>%
  mutate(country = countrycode::countrycode(country, "country.name", "country.name"))
```



## WDI 

* `unempl` = (`SL.UEM.TOTL.ZS`): Unemployment, total (% of total labor force) (modeled ILO estimate)
* `ref_pop` = (`SM.POP.REFG`): Refugee population by country or territory of asylum Search data e.g. GDP, population, Indonesia DataBankMicrodataData Catalog Refugee population by country or territory of asylum.

```{r}
library(WDI)
wdi <- WDI::WDI(
  country = "all", 
  indicator = c("SL.UEM.TOTL.ZS", "SM.POP.REFG"),
  start = 2000, 
  end = 2017
) 

wdi_final <- wdi %>% 
  rename(unempl = SL.UEM.TOTL.ZS, ref_pop = SM.POP.REFG) %>%
  mutate(year = as.character(year)) %>%
  mutate(country = countrycode::countrycode(country, "country.name", "country.name")) %>%
  filter(!is.na(country))

glimpse(wdi_final)
```


## Merge Macro

```{r}
macro_all <- vdem_final %>%
  mutate(year = as.character(year)) %>%
  mutate(country = as.character(country)) %>%
  left_join(hdi_final, by = c("country", "year")) %>%
  left_join(wdi_final, by = c("country", "year"))
glimpse(macro_all)
```


```{r}
sel <- ess_final %>%
  group_by(country, round_year) %>%
  tally

macro_final_list <- list()
for(jj in seq_along(sel$country)){
  temp <- final_macro %>%
    filter(country == sel$country[jj]) %>%
    filter(year %in% (sel$round_year[jj]-2):(sel$round_year[jj]+2))
  
    if(nrow(temp) == 0){
      temp <- final_macro %>%
        filter(country == sel$country[jj]) %>%
        arrange(desc(year)) %>%
        slice(1:3)
    }

  macro_final_list[[jj]] <- temp
}

macro_final <- bind_rows(macro_final_list) %>%
  group_by(country) %>%
  summarise_if(is.double, mean, na.rm = T) %>%
  ungroup() %>%
  select(-e_mipopula)

# save(macro_final, file = "data/Rdata/macro_final.Rdata")
```
