---
title: "CHES Clustering"
subtitle: "KNN + Treebased clustering methods"
author: "Rebecca & Simon "
output: github_document
---

## Packages

```{r}
pacman::p_load(dplyr, ggplot2, readr, haven, broom, purrr, tidyr, magrittr, labelled, sjPlot, viridis, forcats, ggthemes, cluster, factoextra, fpc)
ggplot2::theme_set(ggthemes::theme_few())
```

## Data

```{r}
ches <- get(load("data/Rdata/ches_final.Rdata"))
normalize_range <- function(x){(x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm = T))}
ches <- ches %>%
  mutate(populism = antielite_salience + corrupt_salience) %>% 
  mutate(populism2 = normalize_range(normalize_range(antielite_salience) + 
                               (1 - normalize_range(eu_position)))*100) %>% #+ 
                           #    (1 - range01(eu_budgets)))*100) %>% 
  mutate(liberalism = sociallifestyle + civlib_laworder + galtan) %>% 
  mutate(populism = normalize_range(populism)*100) %>% 
  mutate(liberalism = normalize_range(liberalism)*100) 
```

## CLustering

### Model-Based

The model parameters can be estimated using the Expectation-Maximization (EM) algorithm initialized by hierarchical model-based clustering. Each cluster k is centered at the means μk, with increased density for points near the mean.

```{r cluster_prep}
ches_data <- ches %>% 
  drop_na(eu_position, antielite_salience, sociallifestyle, civlib_laworder, galtan) %>% 
  as.data.frame()

ches_cluster <- ches_data %>% 
  select(antielite_salience, eu_position, civlib_laworder, galtan) %>% 
  purrr::map_df(scale) 
```



```{r cluster_fit}
library(mclust)
mc <- Mclust(ches_cluster) # Model-based-clustering 
summary(mc)
```


```{r cluster_label, fig.cap="Final Cluster Labels", eval = F}
# Classification: plot showing the clustering 

gg_mc_horse <- ches_data %>% 
    mutate(var = factor(mc$classification, levels = 1:3)) %>%
  ggplot(aes(liberalism, populism2)) + 
  geom_point(aes(colour = var, shape = var)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = F, color = "black", linetype = "dashed")+
  #geom_text_repel(aes(liberalism, populism, label = party_cntry)) +
  scale_colour_grey() +
  theme_classic() +
  theme(legend.position = "bottom") +
  ggtitle("Horse Curve", subtitle = "U-Shape Distribution")
```


```{r cluster_val}
# BIC values used for choosing the number of clusters 
gg_mc_bic <- fviz_mclust(mc, "BIC", palette = "grey")
ggsave(gg_mc_bic, filename = "images/gg_mc_bic.pdf", width = 7, height = 4)

gg_mc_cluster <- fviz_mclust(mc, "classification", geom = "point",
pointsize = 1.5, palette = "grey")
ggsave(gg_mc_cluster, filename = "images/gg_mc_cluster.pdf", width = 7, height = 4)

gg_mc_uncert <- fviz_mclust(mc, "uncertainty", palette = "grey") +
  theme(legend.position = "bottom")
#gg_cluster <- gridExtra::grid.arrange(gg_mc_cluster, gg_mc_uncert, ncol = 2)
#ggsave(gg_cluster, filename = "images/gg_cluster.pdf", width = 10, height = 4)
```


```{r}
ches_clust <- ches_data %>% 
    mutate(cluster = case_when(
      mc$classification == 1 ~ "Establishment",
      mc$classification == 2 ~ "Left Populist",
      mc$classification == 3 ~ "Right Populist"
    )
  ) %>%
  select(party_id, cluster) %>% 
  right_join(ches)

#save(ches_clust, file = "data/Rdata/ches_clust.Rdata")
```






```{r}
cluster_means <- ches_clust %>%
  group_by(cluster) %>%
  summarise_all(mean) %>%
  mutate(cluster = NULL) %>%
  t
rn <- rownames(cluster_means)
cluster_means <- cbind(rn, cluster_means) %>% as_tibble()
colnames(cluster_means) <- c("","Cluster 1", "Cluster 2", "Cluster 3")

library(kableExtra)
cluster_means %>% 
  knitr::kable(format = "latex", booktabs = T, linesep = "", caption = "Deliberative Quality Index (2003)") %>%
  column_spec(1, width = "7em", bold=T) %>%
  #column_spec(2, width = "13em") %>%
  #column_spec(3, width = "17em") %>%
  #collapse_rows(columns = 1:3) %>%
  kable_styling(font_size = 10, latex_options = "hold_position") %>%
  add_footnote(c("Steenbergen et al 2003: 27-30"), notation = "number")
```



```{r}
devtools::install_github("ricardo-bion/ggradar", 
                          dependencies=TRUE)

library(ggradar)
ches_data %>% 
  select(antielite_salience, eu_position, civlib_laworder, galtan)

mtcars %>%
     rownames_to_column( var = "group" ) %>%
     mutate_at(vars(-group),funs(rescale)) %>%
     tail(4) %>% select(1:10) -> mtcars_radar

ggradar(mtcars_radar) 
```






# Horse-shoe

```{r}
ches_data %>% 
  ggplot(aes(liberalism, populism, colour = eu_position)) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = F)+
  #geom_text_repel(aes(liberalism, populism, label = party_cntry)) +
  ggthemes::theme_hc() +
  viridis::scale_color_viridis()
```



## Clustering

```{r}
set.seed(2018)
ches_cluster_data <- ches %>% 
  select(party_name, vote_id, liberalism, populism2) %>% 
  drop_na(liberalism, populism2) %>% 
  as.data.frame()

ches_cluster <- ches_cluster_data %>% 
  select(-party_name, -vote_id) %>% 
  purrr::map_df(scale) 
```

```{r, eval = F}
distance <- get_dist(ches_cluster)
fviz_dist(distance, 
 gradient = list(low = "#00AFBB", 
                 mid = "white",
                 high = "#FC4E07"))
```

`kmeans()` function returns a list of components, including:

* `cluster`: A vector of integers (from 1:k) indicating the cluster to which each point is allocated
* `centers`: A matrix of cluster centers (clustqer means)
* `totss`: The total sum of squares (TSS), i.e (xi ≠ x ̄)2. TSS measures the total variance in the data.
* `withinss`: Vector of within-cluster sum of squares, one component per cluster
* `tot.withinss`: Total within-cluster sum of squares, i.e. sum(withinss)
* `betweenss`: The between-cluster sum of squares, i.e. totss ≠ tot.withinss
* `size`: The number of observations in each cluster

```{r}
k4 <- kmeans(ches_cluster, centers = 4, nstart = 25, iter.max = 10)
ggf <- fviz_cluster(k4, data = ches_cluster, show.clust.cent = T, text = "vote_id")
ggf + theme_gdocs()
```

```{r}
ches_cluster_data$k4cluster <- k4$cluster

ches_clust <- ches %>%
  left_join(ches_cluster_data)

# save(ches_clust, file = "data/Rdata/ches_clust.Rdata")
```

```{r, fig.width=10}
glimpse(ches_clust)

ches_trick <- ches_clust %>% mutate(k4cluster = NULL)
ches_clust %>% 
  mutate(k4cluster = ifelse(k4cluster %in% c(1,3), 3, k4cluster)) %>%
  ggplot(aes(liberalism, populism2, colour = k4cluster)) + 
  geom_point(data = ches_trick, color = "gray") +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = F)+
  #geom_text_repel(aes(liberalism, populism, label = party_cntry)) +
  ggthemes::theme_hc() +
  viridis::scale_color_viridis() +
  facet_wrap(~k4cluster, ncol = 4) +
  theme(legend.position = "none")
```




```{r}
fviz_cluster(
  k3, 
  data = ches_cluster,
  palette = c("#2E9FDF", "#00AFBB", "#E7B800"), 
  ellipse.type = "euclid", # Concentration ellipse star.plot = TRUE, # Add segments from centroids to items repel = TRUE, # Avoid label overplotting (slow)
  ggtheme = theme_minimal()
)
```



```{r}
# Elbow method
fviz_nbclust(ches_cluster, kmeans, method = "wss") +
  geom_vline(xintercept = 4, linetype = 2) +
  labs(subtitle = "Elbow method") + 
  theme_gdocs()


# Silhouette method
fviz_nbclust(ches_cluster, kmeans, method = "silhouette") +
  labs(subtitle = "Silhouette method") +
  theme_gdocs()

# Gap statistic
# nboot = 50 to keep the function speedy.
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
fviz_nbclust(ches_cluster, kmeans, nstart = 25, method = "gap_stat", nboot = 50) + 
  labs(subtitle = "Gap statistic method") +
  theme_gdocs()
```

According to these observations, it’s possible to define k = 4 as the optimal number of clusters in the data.


2 dimensions do not need pca

```{r, eval = F}
library(purrr)
res <- purrr::map(2:8, ~ kmeans(ches_cluster, .))
library(ggfortify)
autoplot(res, data = ches_cluster, ncol = 3) + theme(legend.position = "none")
```

normal scatterplots

```{r}
k_cluster_dat <- 2:8 %>%
  purrr::map(~ kmeans(ches_cluster, .x)$cluster) %>%
  reduce(cbind) %>%
  as_tibble() %>%
  set_names(paste0("k", 2:8)) %>%
  cbind(ches_cluster, .)

k_cluster_dat %>%
  gather("k", "value", -liberalism, -populism2) %>%
  ggplot(aes(liberalism, populism2, colour = as.factor(value))) +
  geom_point() +
  facet_wrap(~k) +
  scale_colour_viridis(discrete = T, direction = -1) + 
  theme_hc() +
  theme(legend.position = "none")
```



```{r}
cbind(ches_cluster, cluster = k3$cluster) %>%
  group_by(cluster) %>%
  summarise_all(.funs = list(m = mean, s = sd))
```


## K-Medoids

The most common k-medoids clustering methods is the PAM algorithm (Partitioning Around Medoids, Kaufman & Rousseeuw, 1990).

```{r}
res_pam <- pam(ches_cluster, 3, metric = "euclidean", stand = FALSE)
res_pam$clustering

fviz_nbclust(ches_cluster, pam, method = "silhouette")+
theme_hc()
```

```{r}
km_cluster_dat <- 2:8 %>%
  purrr::map(~ pam(ches_cluster, k = ., metric = "euclidean", stand = FALSE)$clustering) %>%
  reduce(cbind) %>%
  as_tibble() %>%
  set_names(paste0("k", 2:8)) %>%
  cbind(ches_cluster, .)

gg_km <- km_cluster_dat %>%
  gather("k", "value", -liberalism, -populism2) %>%
  ggplot(aes(liberalism, populism2, colour = as.factor(value))) +
  geom_point() +
  facet_wrap(~k) +
  scale_colour_viridis(discrete = T, direction = -1) + 
  theme_hc() +
  theme(legend.position = "none")
gg_km
```

```{r}
res_pam$medoids
```

```{r}
cbind(ches_cluster, cluster = res_pam$clustering) %>%
  group_by(cluster) %>%
  summarise_all(.funs = list(m = median))
```

```{r}
fviz_nbclust(ches_cluster, clara, method = "silhouette")+
theme_classic()
```

# HCA

```{r}
res.dist <- dist(ches_cluster, method = "euclidean")
res.hc <- hclust(d = res.dist, method = "ward.D2")
fviz_dend(res.hc, cex = 0.5)
```

```{r}
# Cut tree into 3 groups
grp <- cutree(res.hc, k = 3)

fviz_dend(
  res.hc, 
  k = 3, # Cut in four groups
  cex = 0.5, # label size
  k_colors = c("#2E9FDF", "#00AFBB", "#E7B800"),
  color_labels_by_k = TRUE, # color labels by groups
  rect = TRUE # Add rectangle around groups
)
```

```{r, fig.height=10, fig.width=10}
fviz_dend(
  res.hc, 
  cex = 1, 
  k = 3,
  k_colors = "jco", 
  type = "circular"
)
```

```{r}
require("igraph")
ggrep <- fviz_dend(res.hc, k = 3, k_colors = "jco",
          type = "phylogenic", repel = TRUE)
ggrep
```

```{r}
fviz_dend(res.hc, k = 3, # Cut in four groups
          k_colors = "jco",
          type = "phylogenic", 
          repel = TRUE,
          phylo_layout = "layout_with_drl")
```


```{r}
fviz_dend(res.hc, k = 3, # Cut in four groups
          k_colors = "jco",
          type = "phylogenic", 
          repel = TRUE,
          phylo_layout = "layout_as_tree")
```


```{r}
fviz_dend(res.hc, k = 3, # Cut in four groups
          k_colors = "jco",
          type = "phylogenic", 
          repel = TRUE,
          phylo_layout = "layout.gem")
```


```{r}
fviz_dend(res.hc, k = 3, # Cut in four groups
          k_colors = "jco",
          type = "phylogenic", 
          repel = TRUE,
          phylo_layout = "layout.mds")
```


```{r}
gg10 <- fviz_dend(res.hc, k = 3, # Cut in four groups
          k_colors = "jco",
          type = "phylogenic", 
          repel = TRUE,
          phylo_layout = "layout_with_lgl")
gg10
```


## Compare clustering algorithms in R

```{r}
library(clValid)
# Iris data set:
# - Remove Species column and scale df <- scale(iris[, -5])
# Compute clValid
clmethods <- c("hierarchical","kmeans","pam") 
intern <- clValid(
  ches_cluster %>% as.matrix, 
  nClust = 2:8,
  clMethods = clmethods, 
  validation = "internal"
) 
summary(intern)
```

```{r, eval = F}
# Stability measures
clmethods <- c("hierarchical","kmeans","pam")
stab <- clValid(
  ches_cluster %>% as.matrix,
  nClust = 2:6, 
  clMethods = clmethods,
  validation = "stability"
) # Display only optimal Scores
optimalScores(stab)
```


## Model-Based Clustering

The model parameters can be estimated using the Expectation-Maximization (EM) algorithm initialized by hierarchical model-based clustering. Each cluster k is centered at the means μk, with increased density for points near the mean.


```{r}
library(mclust)
mc <- Mclust(ches_cluster) # Model-based-clustering 
summary(mc)
```
```{r}
# Classification: plot showing the clustering 
fviz_mclust(mc, "classification", geom = "point",
pointsize = 1.5, palette = "grey") # Classification uncertainty
```


```{r cluster_val, fig.width=12, fig.width=5, fig.cap="Internal Cluster Validation and Uncertainty"}
# BIC values used for choosing the number of clusters 
gg_mc_bic <- fviz_mclust(mc, "BIC", palette = "grey")+
   theme(legend.position = "none")
gg_mc_uncert <- fviz_mclust(mc, "uncertainty", palette = "grey") +
   theme(legend.position = "none")
gg_val <- gridExtra::grid.arrange(gg_mc_bic, gg_mc_uncert, ncol = 2)
ggsave(gg_val, filename = "images/gg_val.png", width = 12, height = 5)
```


```{r}
ches %>% 
  ggplot(aes(liberalism, populism2, colour = mc$classification)) + 
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 2))+
  #geom_text_repel(aes(liberalism, populism, label = party_cntry)) +
  ggthemes::theme_hc() +
  viridis::scale_color_viridis() + 
  geom_density2d(alpha = .7, color = "gray") # Add 2D density 
```



```{r, eval = F}
library(highcharter)

df1 <- tibble(vote = paste(ches_cluster_data$party_name, ches_cluster_data$vote_id, sep = " "), cluster = k3$cluster, x = as.vector(ches_cluster$liberalism), y = as.vector(ches_cluster$populism2)) %>%
  filter(stringr::str_detect(vote, "DE_"))
hchart(df1, hcaes(x = x, y = y, name = vote, color = cluster), type = 'scatter') %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_tooltip(
    formatter = JS("function(){
                    return ('Party: <strong>' + this.point.vote + '</strong><br> X: ' + this.x + ' <br> Y: ' + this.y + ' <br>')
                  }")) %>%
  hc_chart(zoomType = "xy")
```

## World Map

```{r}
library(ggplot2)
world <- map_data("world")
world$iso3 <- countrycode::countrycode(world$region, "country.name", "iso3c")

ches <- ches %>%
  mutate(country = stringr::str_replace(vote_id, "_.*?$", "")) %>%
  mutate(iso3 = countrycode::countrycode(country, "iso2c", "iso3c"))
world$value <- ifelse(world$iso3 %in% unique(ches$iso3), "yes", "no")
# table(world$value)
# world %>% 
#   ggplot(aes(long, lat, group = group)) + 
#     geom_polygon(fill='grey')

ggmap <- world %>% 
  ggplot(aes(long, lat, group = group, fill = value)) + 
  geom_polygon() +
  #xlim(-20,50) + 
  #ylim(30,80) +
  scale_fill_manual("selected", values = c("gray90", "blue")) +
  theme_map()
ggmap
```



```{r, eval = F}
#ess_clean <- ess_sub  %>% 
  # mutate(eu_member =
  #          recode_factor(cntry,
  #               DE = 1958, BE = 1958, FR = 1958, NL = 1958, IE = 1973,
  #               GB = 1973, FI = 1995, AT = 1995, SE = 1995, EE = 2004,
  #               PL = 2004, SI = 2004, CZ = 2004, CH = 0, IL = 0,
  #               IS = 0, NO = 0, RU = 0
  #             )
  #       ) %>%
  # mutate(post_com = ifelse(region %in% c("Estonia", "Poland", "Slovenia", "Czech Republic", "Russian Federation"), "Post C", "West"))
```


